Data Aggregation System - a system for information retrieval on demand over relational and non-relational distributed data sources

### 2011-01-03 doc/chep_paper/das_chep2010_proceedings.tex
Update DAS CHEP paper; re-factor UI to make external links; fixes #880
Signed-off-by: Valentin Kuznetsov <vkuznet@gmail.com>


git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@11242 4525493e-7705-40b1-a816-d608a930855b

\begin{document}
\title{Data Aggregation System - a system for information retrieval on demand over relational and non-relational distributed data sources}
\author{G Ball$^1$, V Kuznetsov$^2$, D Evans$^3$ and S Metson$^4$}

@@

\begin{abstract}
We present the Data Aggregation System, a system for information retrieval and aggregation from heterogenous sources of relational and non-relational data for the Compact Muon Solenoid experiment on the CERN Large Hadron Collider. The experiment currently has a number of organically-developed data sources, including front-ends to a number of different relational databases and non-database data services which do not share common data structures or APIs, and cannot at this stage be readily converged. DAS provides an single interface for querying all these services, a caching layer to speed up access to expensive underlying calls and the ability to merge records from different data services pertaining to a single primary key.
\end{abstract}

@@

In 2010, data taking at the CMS experiment began in earnest, with approximately $42pb^{-1}$ of proton-proton collision data recorded at the time of writing. At full data-taking rates, this is expected to produce around 5PiB of raw and reconstructed data per year. Associated with this is around 1TiB per year of metadata, such as data locations, dataset descriptions and machine conditions.

@@

The architecture of DAS was designed as a series of independent components, which can scale from all running on a single node to multiple nodes and duplicate components, as necessary (Fig.~\ref{dasarch}).

The \emph{web server} handles user sessions, whether they originate from a browser or automated scripts. Queries made here are passed onto the \emph{cache server} for processing, during which time the \emph{web server} periodically polls for the current status (displayed to the user by AJAX) until the request is completed. An MD5 hash of the input query is used to identify the correct output (and hence multiple users making identical queries will only result in a single worker thread doing the processing). Methods are provided here for output in machine-readable formats as well as human-readable formatting.

The \emph{cache server} consists of a pool of worker threads which handle the DAS queries received from the web front-end. Each query is handled entirely by a single thread. A detailed description of the workflow is given in \cite{das2010}. The worker thread first parses a text-based query from the DAS Query Language (Sec.~\ref{dasql}) into an internal representation. The query is then analysed to identify the set of data service APIs which are relevant. The \emph{cache} is checked for items of data matching the query, or data that would form a superset of the query. If any data cannot be found, the relevant service (Sec.~\ref{services}) is called to fetch and transform the data, and insert it into the raw cache (Sec.~\ref{cachemerge}). Once all the data is in the raw cache, records sharing common keys are merged together into a single document, which is stored in the merge cache. If any additional data processing has been specified in the query, it is performed on the merged document before being returned.

The \emph{cache} consists of one or more MongoDB~\cite{mongodb} shards. MongoDB is a document store which natively stores the JSON documents DAS uses as its internal representation. This is used both for the primary record stores (the \emph{raw} and \emph{merged} caches) and for the ancillary databases required by DAS, such as storing the server logs, analytics data and mapping between keys. MongoDB also provides GridFS~\cite{gridfs}, which DAS uses to store otherwise oversize documents, typically the result of merging large numbers of raw documents together.

@@

\section{Benchmarks}

\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{das_vs_mongo_fully_random}
\end{center}
\caption{\label{dasbench}Benchmark showing the average latency for fetching a single randomly-selected document from a wildcard query, with a given number of parallel clients. The cache contains $5\times10^7$ records. The red line shows the performance of the complete DAS stack, and the blue line the time spent by MongoDB.}
\end{figure}

To benchmark DAS performance, we used a 64-bit linux node with 8 cores (each 2.33GhZ) and 16GiB of RAM, which we would expect to be typical of the hardware DAS would use in production. All DAS systems and MongoDB share this node.

@@

\section*{References}
\begin{thebibliography}{9}
\bibitem{drsg} DISCOVER RSG {\it http://drsg.cac.cornell.edu/content/drsg-pilot-projects}
\bibitem{das2010} Kuznetsov V, Evans D and Metson S 2010 The CMS data aggregation system {\it Procedia Comp. Sci.} {\bf 1} 1529-37
\bibitem{mongodb} MongoDB, a scalable, high-performance, open-source, document-orientated database {\it http://www.mongodb.org/}
\bibitem{gridfs} MongoDB GridFS {\it http://www.mongodb.org/display/DOCS/GridFS}
\bibitem{dbsql} Kuznetsov V, Riley D, Afaq A, Sekhri V, Guo Y and Lueking L 2009 The CMS DBS query language {\it CHEP}
\bibitem{ply} Python Lex-Yacc {\it http://www.dabeaz.com/ply/}
\bibitem{yaml} YAML ain't markup language {\it http://www.yaml.org/}
\bibitem{phedex} Rehn J, Barrass T, Bonacorsi D, Hernandez J, Semeniouk I, Tuura L and Wu Y 2006 PhEDEx high-throughput data transfer management system {\it CHEP}
\end{thebibliography}
\end{document}

>>>>

\begin{document}
\title{Data Aggregation System - a system for information retrieval on demand over relational and non-relational distributed data sources}
\author{G. Ball$^1$, V. Kuznetsov$^2$, D. Evans$^3$ and S. Metson$^4$}

@@

\begin{abstract}
We present the Data Aggregation System, a system for information retrieval and aggregation from heterogenous sources of relational and non-relational data for the Compact Muon Solenoid experiment on the CERN Large Hadron Collider. The experiment currently has a number of organically-developed data sources, including front-ends to a number of different relational databases and non-database data services which do not share common data structures or APIs {\it (Application Programming Interfaces)}, and cannot at this stage be readily converged. DAS provides a single interface for querying all these services, a caching layer to speed up access to expensive underlying calls and the ability to merge records from different data services pertaining to a single primary key.
\end{abstract}

@@

In 2010, data taking at the CMS experiment began in earnest, with approximately $42pb^{-1}$ of proton-proton collision data recorded at the time of writing. At full data-taking rates, this is expected to produce around 5PB of raw and reconstructed data per year. Associated with this is around 1TB per year of metadata, such as data locations, dataset descriptions and machine conditions.

@@

The architecture of DAS was designed as a series of independent components, which can scale from all running on a single node to multiple nodes and duplicate components, as necessary (Fig.~\ref{dasarch}). An initial description of the architectur, choice of technology and comparison with similar systems was given in \cite{das2010}. Here we present further development of the query language (Sec.~\ref{dasql}), analytics (Sec.~\ref{analytics}) and initial benchmark results (Sec.~\ref{benchmarks}).

The \emph{web server} handles user sessions, whether they originate from a browser or automated scripts. Queries made here are passed onto the \emph{cache server} for processing, during which time the \emph{web server} periodically polls for the current status and displays it to the user by AJAX {\it (Asynchronous Javascript and XML)}, until the request is completed. An MD5 hash of the input query is used to identify the correct output (and hence multiple users making identical queries will only result in a single worker thread doing the processing). Methods are provided here for output in machine-readable formats as well as human-readable formatting.

The \emph{cache server} consists of a pool of worker threads which handle the DAS queries received from the web front-end. Each query is handled entirely by a single thread. The worker thread first parses a text-based query from the DAS Query Language (Sec.~\ref{dasql}) into an internal representation. The query is then analysed to identify the set of data service APIs which are relevant. The \emph{cache} is checked for items of data matching the query, or data that would form a superset of the query. If any data cannot be found, the relevant service (Sec.~\ref{services}) is called to fetch and transform the data, and insert it into the raw cache (Sec.~\ref{cachemerge}). Once all the data is in the raw cache, records sharing common keys are merged together into a single document, which is stored in the merge cache. If any additional data processing has been specified in the query, it is performed on the merged document before being returned.

The \emph{cache} consists of one or more MongoDB~\cite{mongodb} shards. MongoDB is a document store which natively stores the JSON {\it (Javascript Object Notation)} documents DAS uses as its internal representation. This is used both for the primary record stores (the \emph{raw} and \emph{merged} caches) and for the ancillary databases required by DAS, such as storing the server logs, analytics data and mapping between keys. MongoDB also provides GridFS~\cite{gridfs}, which DAS uses to store otherwise oversize documents, typically the result of merging large numbers of raw documents together. The choice of a document store instead of a traditional relational database was driven by the need to be able to store and subsequently query individual members of the deeply-nested data structures that DAS handles, the structure of which are not known until run-time. We tested the CouchDB~\cite{couchdb} and MongoDB document stores for this purpose; we found the latter to have better performance in our use-case, better scalability across multiple nodes and a query model more compatible with our own.

@@

\section{\label{benchmarks}Benchmarks}

\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{das_vs_mongo_fully_random}
\end{center}
\caption{\label{dasbench}Benchmark showing the average latency for fetching a single randomly-selected document from a wildcard query, with a given number of parallel clients. The cache contains $5\times10^7$ records. The red triangles show the performance of the complete DAS stack, and the blue stars the time spent by MongoDB.}
\end{figure}

To benchmark DAS performance, we used a 64-bit linux node with 8 cores (each 2.33GhZ) and 16GB of RAM, which we would expect to be typical of the hardware DAS would use in production. All DAS systems and MongoDB share this node.

@@

\section*{References}
\begin{thebibliography}{9}
\bibitem{drsg} DISCOVER RSG {\it http://drsg.cac.cornell.edu/content/drsg-pilot-projects}
\bibitem{das2010} Kuznetsov V, Evans D and Metson S 2010 The CMS data aggregation system {\it Procedia Comp. Sci.} {\bf 1} 1529-37
\bibitem{mongodb} MongoDB, a scalable, high-performance, open-source, document-orientated database {\it http://www.mongodb.org/}
\bibitem{gridfs} MongoDB GridFS {\it http://www.mongodb.org/display/DOCS/GridFS}
\bibitem{couchdb} Apache CouchDB {\it http://couchdb.apache.org/}
\bibitem{dbsql} Kuznetsov V, Riley D, Afaq A, Sekhri V, Guo Y and Lueking L 2009 The CMS DBS query language {\it CHEP}
\bibitem{ply} Python Lex-Yacc {\it http://www.dabeaz.com/ply/}
\bibitem{yaml} YAML ain't markup language {\it http://www.yaml.org/}
\bibitem{phedex} Rehn J, Barrass T, Bonacorsi D, Hernandez J, Semeniouk I, Tuura L and Wu Y 2006 PhEDEx high-throughput data transfer management system {\it CHEP}
\end{thebibliography}
\end{document}
